---
title: "Regression Models Course Project"
author: "André Marinho"
date: "24/08/2020"
output:
        html_document:
                keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 0. Executive Summary

The Motor Trend US magazine is a very famous magazine in the United States for producing accurate data about automobile. The Motor Trend Car Road Tests (mtcars) is a dataset that was extracted from the 1974 and comprises fuel consumption based on 10 aspects of 32 models of automobile design and performance, collected by the Motor Trend US magazine.

This project analysis mtcars with a particular interest in the following questions:

- Is an automatic or manual transmission better for MPG?
- Quantify the MPG difference between automatic and manual transmissions.

It is conclude that there is a significant difference between the MPG of manual and automatic transmission vehicles. The manual automobiles have a better fuel efficiency (2.94 MPG more) than automatic ones.

## 1. Exploratory Data Analysis

```{r message=FALSE}

# Load libraries
library(ggplot2)
library(dplyr)
library(corrgram)
library(caret)

# Data analyses
data <- mtcars
dim(data)
head(data)
summary(data)

# Data correlation
corr <- cor(data)
corr_high <- findCorrelation(corr, cutoff=0.75)
names(data)[corr_high]

# t-test
t_test <- t.test(mpg~am, data=data)
t_test$p.value
t_test$estimate

# summary of mpg, by transmission type
data %>%
        mutate(tr_type = ifelse(am == 0, no="manual", yes="automatic")) %>%
        group_by(tr_type) %>%
        summarise(median=median(mpg), mean=mean(mpg), sd=sd(mpg), min=min(mpg), max=max(mpg), count=n()) 
```

Data analyses conclusions:

1. We can see higher correlations between "cyl", "disp", "mpg", and "gear".

2. The p-value is less than 0.05. The null hypothesis "there is no correlation between transmission type and mpg" is rejected.

3. There is more "automatic" data (6 more), but their median, mean, standard deviation, min, and max are all less than corresponding "manual" values.

**Note:** refer to the **Appendix** section for the plotting.

## 2. Build Model

### 2.1. Linear Regression

```{r}
linear_model <- lm(mpg~am, data=data)
summary(linear_model)$adj.r.squared
```

The value of ~0.34 represents ~34% (a low value) of the variation in the linear model. Hence, we going to examine other models. 

### 2.2. Multi-variate Model

```{r}
model <- lm(mpg~., data=data)
step_model <- step(model, k=log(nrow(data)), trace=FALSE)
summary(step_model)$adj.r.squared
summary(step_model)
```

Based on the step function, the optimal set of features is: {mpg ~ wt + qsec + am}. The variation is ~83% (an high value) in this case. Refer to the **Appendix** section for the plotting.

### 2.3. Fit and Select Model

```{r}
best_model <- lm(mpg ~ wt + qsec + am, data=data)
summary(best_model)$adj.r.squared # confirming adj.r.squared

# Select final model
anova(linear_model, step_model, best_model)
confint(best_model)
summary(best_model)$coef
```

Refer to **Appendix** section for the residual plotting.

## 3. Conclusions

Based on the models, plotting, and adopted strategy, we can conclude:

1. The multivariate model suggests an am of 2.94 MPG for Manual (better) transmission over Automatic.

2. The weight of a car holds to the fuel efficiency of lighter vehicles.

3. Residual plotting provides heteroscedasticity information of the built models.

4. The Residuals vs Fitted plots shows the difference about the independence of the variables.

5. The Normal Q-Q plots displays a normal distribution of the points.

6. The Scale-Location plots shows the variance of the points randomly distributed.

7. The Residuals vs Leverage plots displays that values fall within 0.5 mark.

## 4. Appendix

### 4.1. Violin Plot (Data Analysis)

```{r}
data_plot <- data
data_plot %>%
        mutate(am = ifelse(am == 0, no='manual', yes='automatic')) %>%
        ggplot(aes(x=am, y=mpg)) +
        geom_violin(aes(fill=factor(am))) +
        geom_jitter(height=0)
```

### 4.2. Motor Trend Card Road Tests Data Correlogram

```{r}
corrgram(data, order=TRUE, upper.panel=NULL, main="Car Data Correlogram", lower.panel=panel.cor)
```

### 4.3. Optimal Variables Plotting

```{r}
data_optimal <- select(data, mpg, wt, qsec, am)
pairs(x=data_optimal, panel=panel.smooth, main="Optimal Variables Pair Plotting")
```

### 4.4. Residual Plotting

#### 4.4.1. First Model

```{r}
par(mfrow=c(2,2))
plot(linear_model)
```

#### 4.4.2. Better Model

```{r}
par(mfrow=c(2,2))
plot(best_model)
```


